Lattice Boltzmann simulations of the dynamics of thin liquid films \\ \large{Gitter Boltzmann Simulationen der Dynamik von d\"unnen Fl\"ussigkeitsfilmen}

### 2024-01-21 chapters/Paper_two.tex
Updates from Overleaf

\subsection{Testing the stochastic term}\label{subsec:validation}
\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{graphics/spectrum_theta_20_nob_fill_pattern.pdf}
    \caption{Height fluctuations spectra from deterministic (circles) and stochastic (triangles) simulations at $t=0.2 t_0$ (filled symbols) and $t=0.7 t_0$ (empty symbols), on a substrate with $\theta =\pi/9$. 
    The theoretical predictions, Eq.~(\ref{eq:structure_factor_2}), are reported with solid ($\sigma=0$) and dashed ($\sigma = \sigma_0$) lines.}
    \label{fig:theory_simulation_structure_factor}
\end{figure}
To validate our model against the analytical results discussed in Sec.~\ref{sec:stoch_thin_film}, we first rewrite Eq.~(\ref{eq:structure_factor_2}) in a form, involving the dimensionless temperature $\sigma$, which is more convenient for the forthcoming discussion, namely
\begin{equation}\label{eq:structure_factor_2}
    S(q,t) = S_0 e^{2\omega(q)t} + \frac{\sigma^2 h_0^2 L}{q_0^2}g\left(\frac{q}{q_0}\right)(e^{2\omega(q)t} - 1),
\end{equation}
where $g(\xi) = \frac{\xi^2}{2\xi^2 - \xi^4}$ and the appearance of the system size $L$ stems from the noise amplitude, $\langle |\tilde{\mathcal{N}}(q,t)|^2 \rangle = L$, resulting from discrete Fourier-transforming over a finite length~\cite{zhangMolecularSimulationThin2019,zhangNanoscaleThinfilmFlows2020}. 
To arrive at Eq.~(\ref{eq:structure_factor_2}), use was made of Eq.~(\ref{eq:t0}) for the characteristic time $t_0$ and we omitted the dependence on $q$ of $S_0(q)$ because, with our initialization, the latter is just a constant, $S_0(q)\equiv S_0$.
We measured, then, the spectra $S(q,t)$, at two instants of time, in the early stages of the growth of the instability, for both athermal and fluctuating dewetting on a substrate with $\theta=\pi/9$ (and strict no-slip, $\delta=0$).
The data, reported in Fig.~\ref{fig:theory_simulation_structure_factor}, from both deterministic (circles) and fluctuating (triangles) simulations show good agreement with the theoretical curves, Eq.~(\ref{eq:structure_factor_2}), depicted with solid ($\sigma=0$) and dashed ($\sigma=\sigma_0$) lines.
In particular, notice that the maximum of $S(q,t)$ is attained for $q=q_0$ independently of time, in the athermal case, and at $q = q_m > q_0$, with $q_m$ tending to $q_0$ as the time goes by, when fluctuations are present; in the latter case, in addition, the capillary wave spectrum $S_{\text{cw}}(q) \propto q^{-2}$ is recovered for large $q$.

\subsection{Droplet size distributions}\label{subsec:morphandrup}
\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{graphics/Droplet_height_['00', '1e-7']_distri_new_35_nodist.pdf}
    \caption{Histograms of the droplet height distributions from the deterministic (blue, line-patterned) and stochastic (orange, dot-patterned) simulations with $\theta = \pi/9$, at $t \approx 20 t_0$; the corresponding Gaussian kernel density estimations are also plotted, as a guide to the eye, with solid blue (dashed orange) curve for the deterministic (stochastic) data.}
    \label{fig:droplet_distribution}
\end{figure}
In the early stage of the deterministic dewetting, we expect that the droplet size distribution, quantified in terms of the droplet height, will be strongly correlated to the maximum unstable wavelength $\lambda_{\text{max}} =2\pi/q_0$:
\begin{equation}\label{eq:lambda_max}
    \lambda_{\text{max}} = \frac{2\pi}{q_0} = \sqrt{\frac{8\pi^2\gamma}{\Pi'(h)|_{h_0}}} 
\end{equation}
More quantitatively, in our one-dimensional case the equilibrium shape of a droplet is a circular arc, whose chord is the portion of droplet in contact with the substrate and equals $\lambda_{\text{max}}/2$. 
Therefore, it can be shown, by means of simple geometrical arguments, that the droplet height is
\begin{equation}\label{eq:cap_height}
   h_{\text{drop}} = \frac{\lambda_{\text{max}}}{4}\tan\left(\frac{\theta}{2}\right), 
\end{equation}
where $\theta$ is the contact angle. 
For $\theta = \pi/9$, we get $\lambda_{\text{max}} \approx 62$ and $h_{\text{drop}} \approx 2.7$, which is the maximum around which the height values should be distributed.
This is, indeed, what we observe in Fig.~\ref{fig:droplet_distribution}, where we show measurements of the droplet height distribution obtained from simulations with and without thermal fluctuations.
In the deterministic case ($\sigma=0$), the distribution (blue, line-patterned, histogram) sharply peaks around the theoretically predicted value $h_{\text{drop}} \approx 2.7$. 
As a guide to the eye we add a Gaussian kernel density estimation (solid blue curve) that shows the best fitting continuous distribution to the histogram. 
The data from the stochastic simulations ($\sigma = \sigma_0$), instead, show a broadening of the distribution (orange, dot-patterned, histogram and dashed orange curve).
Adding thermal energy, then, on the one hand, facilitates the system to explore higher energetic states, whereas on the other hand it reduces the coarsening time scales~\cite{grunThinFilmFlowInfluenced2006}: this is reflected in the tails of the distribution for both small and large values of $h$, respectively.
These findings are in agreement with what was reported by Nesic \textit{et al.}~\cite{nesicFullyNonlinearDynamics2015}.
\begin{figure}
    \centering
    \includegraphics[width = 0.75\textwidth]{graphics/Correct_t0_normed_delta_h_evo_seminar.pdf}
    \caption{Time evolution of $\Delta h(t) = \max_x\{h(x,t)\} - \min_x\{h(x,t)\}$ for the athermal (bullets) and fluctuating (triangles) systems with $\theta=\pi/9$.
    The blue dashed-dotted line depicts an exponential fit of the data for the athermal case ($\sigma=0$).
    The vertical lines mark the rupture times for the deteministic (solid) and stochastic (dotted) simulations.}
    \label{fig:delta_heigth_evo}
\end{figure}

@@

    \includegraphics[width=0.75\textwidth]{graphics/Andrea_model_t0_normed_rupture_times.pdf}

@@

    \includegraphics[width=0.75\textwidth]{graphics/evolution_qm_with_inset_slip0.pdf}

@@

    \includegraphics[width=0.95\textwidth]{graphics/spacedepCA_['sine', '1e7', '10', 25, '9_3', 10000000.0]_v2.png}

@@

    \includegraphics[width=0.95\textwidth]{graphics/psd_spacedepCA_sine_25_10_00_[0, 450000, 1150000]_9_3_rescaled.pdf}

@@

    \includegraphics[width=0.95\textwidth]{graphics/psd_spacedepCA_sine_25_10_1e-7_[0, 450000, 1150000]_9_3_rescaled.pdf}

@@

\subsubsection{Square wave pattern}\label{subsubsec:square_wave}
\begin{figure}
    \centering
    \includegraphics[width=0.95\textwidth]{graphics/spacedepCA_['delta', '1e7', '10', 25, '9_3', 2000000.0]_v2.png}
    \caption{Space-time plot of the height field $h(x,t)$ evolution over a square-wave patterned substrate undergoing athermal (a) and fluctuating (b) dewetting, respectively. 
    In panel (c) we report the contact angle profile $\theta(x)$ [Eq.~(\ref{eq:sharp_contact_angle_spatial})].} 
    \label{fig:patterned_step8_difference_20-30}
\end{figure}
In the following we focus on the dewetting process on a substrate with a contact angle pattern given by 
\begin{equation}\label{eq:sharp_contact_angle_spatial}
    {\theta^{(2)}(x)} = 
    \left\{
    \begin{array}{ll}
        \frac{\pi}{9} & \text{if} \quad\sin(q_{\theta}x) \leq 0, \\
        \frac{\pi}{6} & \text{otherwise}
    \end{array}
    \right..
\end{equation}
Looking at Fig.~\ref{fig:patterned_step8_difference_20-30}, one can immediately notice two main differences in comparison to the sinusoidal pattern (cf.~Fig.~\ref{fig:patterned_sine8_difference_20-30}). 
The first one is that the impact of thermal fluctuations on the global dewetting morphology looks much weaker; in particular, in both deterministic and stochastic simulations, the film ruptures exactly at the wettability discontinuities.
Second, stable droplets are formed also in regions of high contact angle. 
As a consequence, a total amount of $2\frac{L}{\lambda_{\theta}}$ droplets are observed, which is twice as many as for the pattern $\theta^{(1)}(x)$.

The characteristic timescale of formation process, however, is not the same for all droplets, since it depends on the local contact angle.
We observe, indeed, that droplets nucleate faster in regions of higher contact angle. 
Specifically, if we define the droplet formation time as the delay between the $\theta$-dependent droplet nucleation time, $t_d(\theta)$, and the film rupture time, $\tau_r(\theta)$, i.e.,
\begin{equation}\label{eq:taud}
  \tau_d(\theta) = t_d(\theta) - \tau_r(\theta), 
\end{equation}
and measure the ratio
\begin{equation}\label{eq:time_ratio_delta8_substrate}
    \Xi = \frac{\tau_d(\pi/6)}{\tau_d(\pi/9)} 
\end{equation}
of formation times over patches of different contact angle, then we find $\Xi \approx 0.3$.
Theoretically, this can be explained according to a retracting film scenario. 
We expect, in fact, that the characteristic time Eq.~(\ref{eq:taud}) will be inversely proportional to the speed of the receding contact line over a particular substrate patch, namely, $t_d \propto U^{-1}$; such speed, in turn, is known to depend on the contact angle as $U(\theta) \sim \theta^3$~\cite{snoeijerAsymptoticAnalysisDewetting2010}; therefore, we get for the ratio $\Xi$:
\begin{equation}\label{eq:ratio_U_theta_qubed_1/3}
  \Xi = \frac{\tau_d(\pi/6)}{\tau_d(\pi/9)}  \propto  {\frac{U(\pi/9)}{U(\pi/6)} \sim \frac{\left(\pi/9\right)^3}{\left(\pi/6\right)^3} = \left(\frac{2}{3}\right)^3 = 0.296...},
\end{equation}
in excellent agreement with the measured value.
\begin{figure}
    \centering
    \includegraphics[width=0.95\textwidth]{graphics/Rupture_events_with_film.pdf}
    \caption{(a), (b) Space-time plots showing the evolution of the height field for athermal (a) and fluctuating (b) dewetting over the square-wave patterned substrate, in a neighbourhood of the instant of time at which the first rupture event (in the athermal case) occurred; for the sake of visualization we mark the rupture events with red bullets (\textcolor{red}{$\bullet$}).
    (c) Distribution of times of occurrence of rupture events for the athermal ($\sigma=0$, blue, line-patterned) and fluctuating ($\sigma = \sigma_0$, orange, dot-patterned) dewetting.} 
    \label{fig:rupture_time_distri_square_wave8}
\end{figure}

@@

 \begin{figure}
     \centering
     \includegraphics[width=0.75\textwidth]{graphics/square_wave_det_stoch.pdf}
     \caption{Spectra of the dewetting film on the $\theta^{(2)}(x)$ pattern, Eq.(\ref{eq:sharp_contact_angle_spatial}), at $t = 0.1t_0(\pi/6)$, with ($\sigma = \sigma_0$, \textcolor{pyorange}{$\blacktriangle$}) and without ($\sigma =0$, \textcolor{pyblue}{$\bullet$}) thermal fluctuations. 
     The vertical lines indicate the substrate wave mode $q_{\theta}$ (solid) and its multiples $2q_{\theta}$ (dashed), $3q_{\theta}$ (thin dashed-dotted) and $7q_{\theta}$ (thick dashed-dotted).}
     \label{fig:square_wave_both}
 \end{figure}
As a final analysis we discuss the short time (pre-rupture) spectra on this substrate as well (see Fig.~\ref{fig:square_wave_both}).
We show the deterministic (blue dashed line with bullets) and stochastic (orange dashed line with triangles) data at the same time $t=0.1\,t_0$.
Under the light of what obtained with the sine wave pattern and given the dewetting morphology shown in Fig.~\ref{fig:patterned_step8_difference_20-30}, we would guess to find the absolute maximum of the spectrum at $q=2 q_{\theta}$.
Surprisingly, instead, the figure shows that, although a local maximum located at $2q_{\theta}$ can indeed be detected, there are several higher peaks at $n q_{\theta}$, with a global maximum at $7q_{\theta} \approx 1.2 q_0$; this probably suggests that the early time dewetting on this kind of substrate is still reminiscent of the spinodal background. 
Nevertheless, the strong bond of the dynamics with the pattern is evident and it is further pointed out by the unexpected observation that corresponding maxima from the deterministic and stochastic spectra are all of the same magnitude.

>>>>

\subsection{Testing the stochastic term}\label{subsec:validation}
\begin{figure}
    \centering
    \includegraphics[width=0.65\textwidth]{graphics/spectrum_theta_20_nob_fill_pattern.pdf}
    \caption{Height fluctuations spectra from deterministic (circles) and stochastic (triangles) simulations at $t=0.2 t_0$ (filled symbols) and $t=0.7 t_0$ (empty symbols), on a substrate with $\theta =\pi/9$. 
    The theoretical predictions, Eq.~(\ref{eq:structure_factor_2}), are reported with solid ($\sigma=0$) and dashed ($\sigma = \sigma_0$) lines.}
    \label{fig:theory_simulation_structure_factor}
\end{figure}
To validate our model against the analytical results discussed in Sec.~\ref{sec:stoch_thin_film}, we first rewrite Eq.~(\ref{eq:structure_factor_2}) in a form, involving the dimensionless temperature $\sigma$, which is more convenient for the forthcoming discussion, namely
\begin{equation}\label{eq:structure_factor_2}
    S(q,t) = S_0 e^{2\omega(q)t} + \frac{\sigma^2 h_0^2 L}{q_0^2}g\left(\frac{q}{q_0}\right)(e^{2\omega(q)t} - 1),
\end{equation}
where $g(\xi) = \frac{\xi^2}{2\xi^2 - \xi^4}$ and the appearance of the system size $L$ stems from the noise amplitude, $\langle |\tilde{\mathcal{N}}(q,t)|^2 \rangle = L$, resulting from discrete Fourier-transforming over a finite length~\cite{zhangMolecularSimulationThin2019,zhangNanoscaleThinfilmFlows2020}. 
To arrive at Eq.~(\ref{eq:structure_factor_2}), use was made of Eq.~(\ref{eq:t0}) for the characteristic time $t_0$ and we omitted the dependence on $q$ of $S_0(q)$ because, with our initialization, the latter is just a constant, $S_0(q)\equiv S_0$.
We measured, then, the spectra $S(q,t)$, at two instants of time, in the early stages of the growth of the instability, for both athermal and fluctuating dewetting on a substrate with $\theta=\pi/9$ (and strict no-slip, $\delta=0$).
The data, reported in Fig.~\ref{fig:theory_simulation_structure_factor}, from both deterministic (circles) and fluctuating (triangles) simulations show good agreement with the theoretical curves, Eq.~(\ref{eq:structure_factor_2}), depicted with solid ($\sigma=0$) and dashed ($\sigma=\sigma_0$) lines.
In particular, notice that the maximum of $S(q,t)$ is attained for $q=q_0$ independently of time, in the athermal case, and at $q = q_m > q_0$, with $q_m$ tending to $q_0$ as the time goes by, when fluctuations are present; in the latter case, in addition, the capillary wave spectrum $S_{\text{cw}}(q) \propto q^{-2}$ is recovered for large $q$.

\subsection{Droplet size distributions}\label{subsec:morphandrup}
\begin{figure}
    \centering
    \includegraphics[width=0.65\textwidth]{graphics/Droplet_height_['00', '1e-7']_distri_new_35_nodist.pdf}
    \caption{Histograms of the droplet height distributions from the deterministic (blue, line-patterned) and stochastic (orange, dot-patterned) simulations with $\theta = \pi/9$, at $t \approx 20 t_0$; the corresponding Gaussian kernel density estimations are also plotted, as a guide to the eye, with solid blue (dashed orange) curve for the deterministic (stochastic) data.}
    \label{fig:droplet_distribution}
\end{figure}
In the early stage of the deterministic dewetting, we expect that the droplet size distribution, quantified in terms of the droplet height, will be strongly correlated to the maximum unstable wavelength $\lambda_{\text{max}} =2\pi/q_0$:
\begin{equation}\label{eq:lambda_max}
    \lambda_{\text{max}} = \frac{2\pi}{q_0} = \sqrt{\frac{8\pi^2\gamma}{\Pi'(h)|_{h_0}}} 
\end{equation}
More quantitatively, in our one-dimensional case the equilibrium shape of a droplet is a circular arc, whose chord is the portion of droplet in contact with the substrate and equals $\lambda_{\text{max}}/2$. 
Therefore, it can be shown, by means of simple geometrical arguments, that the droplet height is
\begin{equation}\label{eq:cap_height}
   h_{\text{drop}} = \frac{\lambda_{\text{max}}}{4}\tan\left(\frac{\theta}{2}\right), 
\end{equation}
where $\theta$ is the contact angle. 
For $\theta = \pi/9$, we get $\lambda_{\text{max}} \approx 62$ and $h_{\text{drop}} \approx 2.7$, which is the maximum around which the height values should be distributed.
This is, indeed, what we observe in Fig.~\ref{fig:droplet_distribution}, where we show measurements of the droplet height distribution obtained from simulations with and without thermal fluctuations.
In the deterministic case ($\sigma=0$), the distribution (blue, line-patterned, histogram) sharply peaks around the theoretically predicted value $h_{\text{drop}} \approx 2.7$. 
As a guide to the eye we add a Gaussian kernel density estimation (solid blue curve) that shows the best fitting continuous distribution to the histogram. 
The data from the stochastic simulations ($\sigma = \sigma_0$), instead, show a broadening of the distribution (orange, dot-patterned, histogram and dashed orange curve).
Adding thermal energy, then, on the one hand, facilitates the system to explore higher energetic states, whereas on the other hand it reduces the coarsening time scales~\cite{grunThinFilmFlowInfluenced2006}: this is reflected in the tails of the distribution for both small and large values of $h$, respectively.
These findings are in agreement with what was reported by Nesic \textit{et al.}~\cite{nesicFullyNonlinearDynamics2015}.
\begin{figure}
    \centering
    \includegraphics[width = 0.65\textwidth]{graphics/Correct_t0_normed_delta_h_evo_seminar.pdf}
    \caption{Time evolution of $\Delta h(t) = \max_x\{h(x,t)\} - \min_x\{h(x,t)\}$ for the athermal (bullets) and fluctuating (triangles) systems with $\theta=\pi/9$.
    The blue dashed-dotted line depicts an exponential fit of the data for the athermal case ($\sigma=0$).
    The vertical lines mark the rupture times for the deteministic (solid) and stochastic (dotted) simulations.}
    \label{fig:delta_heigth_evo}
\end{figure}

@@

    \includegraphics[width=0.65\textwidth]{graphics/Andrea_model_t0_normed_rupture_times.pdf}

@@

    \includegraphics[width=0.65\textwidth]{graphics/evolution_qm_with_inset_slip0.pdf}

@@

    \includegraphics[width=0.75\textwidth]{graphics/spacedepCA_['sine', '1e7', '10', 25, '9_3', 10000000.0]_v2.png}

@@

    \includegraphics[width=0.75\textwidth]{graphics/psd_spacedepCA_sine_25_10_00_[0, 450000, 1150000]_9_3_rescaled.pdf}

@@

    \includegraphics[width=0.75\textwidth]{graphics/psd_spacedepCA_sine_25_10_1e-7_[0, 450000, 1150000]_9_3_rescaled.pdf}

@@

\subsubsection{Square wave pattern}\label{subsubsec:square_wave}
\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{graphics/spacedepCA_['delta', '1e7', '10', 25, '9_3', 2000000.0]_v2.png}
    \caption{Space-time plot of the height field $h(x,t)$ evolution over a square-wave patterned substrate undergoing athermal (a) and fluctuating (b) dewetting, respectively. 
    In panel (c) we report the contact angle profile $\theta(x)$ [Eq.~(\ref{eq:sharp_contact_angle_spatial})].} 
    \label{fig:patterned_step8_difference_20-30}
\end{figure}
In the following we focus on the dewetting process on a substrate with a contact angle pattern given by 
\begin{equation}\label{eq:sharp_contact_angle_spatial}
    {\theta^{(2)}(x)} = 
    \left\{
    \begin{array}{ll}
        \frac{\pi}{9} & \text{if} \quad\sin(q_{\theta}x) \leq 0, \\
        \frac{\pi}{6} & \text{otherwise}
    \end{array}
    \right..
\end{equation}
Looking at Fig.~\ref{fig:patterned_step8_difference_20-30}, one can immediately notice two main differences in comparison to the sinusoidal pattern (cf.~Fig.~\ref{fig:patterned_sine8_difference_20-30}). 
The first one is that the impact of thermal fluctuations on the global dewetting morphology looks much weaker; in particular, in both deterministic and stochastic simulations, the film ruptures exactly at the wettability discontinuities.
Second, stable droplets are formed also in regions of high contact angle. 
As a consequence, a total amount of $2\frac{L}{\lambda_{\theta}}$ droplets are observed, which is twice as many as for the pattern $\theta^{(1)}(x)$.

The characteristic timescale of formation process, however, is not the same for all droplets, since it depends on the local contact angle.
We observe, indeed, that droplets nucleate faster in regions of higher contact angle. 
Specifically, if we define the droplet formation time as the delay between the $\theta$-dependent droplet nucleation time, $t_d(\theta)$, and the film rupture time, $\tau_r(\theta)$, i.e.,
\begin{equation}\label{eq:taud}
  \tau_d(\theta) = t_d(\theta) - \tau_r(\theta), 
\end{equation}
and measure the ratio
\begin{equation}\label{eq:time_ratio_delta8_substrate}
    \Xi = \frac{\tau_d(\pi/6)}{\tau_d(\pi/9)} 
\end{equation}
of formation times over patches of different contact angle, then we find $\Xi \approx 0.3$.
Theoretically, this can be explained according to a retracting film scenario. 
We expect, in fact, that the characteristic time Eq.~(\ref{eq:taud}) will be inversely proportional to the speed of the receding contact line over a particular substrate patch, namely, $t_d \propto U^{-1}$; such speed, in turn, is known to depend on the contact angle as $U(\theta) \sim \theta^3$~\cite{snoeijerAsymptoticAnalysisDewetting2010}; therefore, we get for the ratio $\Xi$:
\begin{equation}\label{eq:ratio_U_theta_qubed_1/3}
  \Xi = \frac{\tau_d(\pi/6)}{\tau_d(\pi/9)}  \propto  {\frac{U(\pi/9)}{U(\pi/6)} \sim \frac{\left(\pi/9\right)^3}{\left(\pi/6\right)^3} = \left(\frac{2}{3}\right)^3 = 0.296...},
\end{equation}
in excellent agreement with the measured value.
\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{graphics/Rupture_events_with_film.pdf}
    \caption{(a), (b) Space-time plots showing the evolution of the height field for athermal (a) and fluctuating (b) dewetting over the square-wave patterned substrate, in a neighbourhood of the instant of time at which the first rupture event (in the athermal case) occurred; for the sake of visualization we mark the rupture events with red bullets (\textcolor{red}{$\bullet$}).
    (c) Distribution of times of occurrence of rupture events for the athermal ($\sigma=0$, blue, line-patterned) and fluctuating ($\sigma = \sigma_0$, orange, dot-patterned) dewetting.} 
    \label{fig:rupture_time_distri_square_wave8}
\end{figure}

@@

 \begin{figure}
     \centering
     \includegraphics[width=0.65\textwidth]{graphics/square_wave_det_stoch.pdf}
     \caption{Spectra of the dewetting film on the $\theta^{(2)}(x)$ pattern, Eq.(\ref{eq:sharp_contact_angle_spatial}), at $t = 0.1t_0(\pi/6)$, with ($\sigma = \sigma_0$, \textcolor{pyorange}{$\blacktriangle$}) and without ($\sigma =0$, \textcolor{pyblue}{$\bullet$}) thermal fluctuations. 
     The vertical lines indicate the substrate wave mode $q_{\theta}$ (solid) and its multiples $2q_{\theta}$ (dashed), $3q_{\theta}$ (thin dashed-dotted) and $7q_{\theta}$ (thick dashed-dotted).}
     \label{fig:square_wave_both}
 \end{figure}
As a final analysis we discuss the short time (pre-rupture) spectra on this substrate as well (see Fig.~\ref{fig:square_wave_both}).
We show the deterministic (blue dashed line with bullets) and stochastic (orange dashed line with triangles) data at the same time $t=0.1\,t_0$.
Under the light of what obtained with the sine wave pattern and given the dewetting morphology shown in Fig.~\ref{fig:patterned_step8_difference_20-30}, we would guess to find the absolute maximum of the spectrum at $q=2 q_{\theta}$.
Surprisingly, instead, the figure shows that, although a local maximum located at $2q_{\theta}$ can indeed be detected, there are several higher peaks at $n q_{\theta}$, with a global maximum at $7q_{\theta} \approx 1.2 q_0$; this probably suggests that the early time dewetting on this kind of substrate is still reminiscent of the spinodal background. 
Nevertheless, the strong bond of the dynamics with the pattern is evident and it is further pointed out by the unexpected observation that corresponding maxima from the deterministic and stochastic spectra are all of the same magnitude.
